### 2012-01-21 writeups/proposal/proposal.tex
ported rl_detection writeups in

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\pagestyle{headings}
\usepackage[top=1in, bottom=1in]{geometry}
\usepackage{amsmath,amsfonts,amsthm,amssymb}
\usepackage[pdftex]{graphicx}
\usepackage{url}
\usepackage{subfig}
\usepackage{color}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false,citecolor=red,linkcolor=blue]{hyperref}

%% Command definitions
\def\subsectionautorefname{section}
\newcommand{\note}[1]{\textcolor{red}{\textbf{#1}}}
\definecolor{light-gray}{gray}{0.3}
\newcommand{\aside}[1]{\textcolor{light-gray}{\emph{#1}}}
\newcommand{\comment}[1]{}

\title{Attentional Object Detection: A Proposal}
\author{Sergey Karayev}
\date{updated: 22 April 2011}

\begin{document}
\maketitle

\begin{abstract}
This document tracks the development of ideas for efficient object detection using the idea of attention: a sequential process of looking for something somewhere.
We follow the Deformable Part Model approach to object detection and extend it in several ways, mostly focusing on the role of context, toward the goal of Anytime detection performance.
I expect a submission to NIPS'11 to be whittled down from the text and results here.
\end{abstract}

\tableofcontents
\newpage

\input{introduction.tex}
\input{related_work.tex}
\input{evaluation.tex}
\input{our_approach.tex}
\input{discretization_and_featurization.tex}
\input{ideas.tex}
\input{results.tex}

\bibliographystyle{ieeetr}
\small
\bibliography{../sergeyk-bibtex.bib}
\end{document}