### 2012-01-21 writeups/rl_class/report.tex

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\pagestyle{headings}
\usepackage[top=1in, bottom=1in]{geometry}
\usepackage{amsmath,amsfonts,amsthm,amssymb}
\usepackage[pdftex]{graphicx}
\usepackage{url}
\usepackage{subfig}
\usepackage{color}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false,citecolor=red,linkcolor=blue]{hyperref}

%% Command definitions
\def\subsectionautorefname{section}
\newcommand{\note}[1]{\textcolor{red}{\textbf{#1}}}
\definecolor{light-gray}{gray}{0.3}
\newcommand{\aside}[1]{\textcolor{light-gray}{\emph{#1}}}
\newcommand{\comment}[1]{}

\title{Attentional Object Detection\\
Report for CS 294-67 Sequential Decisions}
\author{Sergey Karayev}
\date{13 May 2011}

\begin{document}
\maketitle

\begin{abstract}
  Multi-class object detection in cluttered natural scenes is a long-standing open problem in computer vision. Currently best-performing approaches are computationally expensive. We seek to ameliorate this through replacing exhaustive and unprioritized search with a sequential decision process.
\end{abstract}

\input{introduction.tex}
\input{related_work.tex}
\input{evaluation.tex}
\input{our_approach.tex}
\input{experiments.tex}

\bibliographystyle{ieeetr}
\small
\bibliography{../sergeyk-bibtex.bib}
\end{document}