### 2012-03-05 writeups/eccv12/evaluation.tex

\section{Evaluation} \label{sec:evaluation}

\todo{Describe random, oracle, and then go through our experimental settings.}

\begin{figure}[h!]
\centering
\subfloat[Detection]{\includegraphics[width=0.78\linewidth]
    {figures/manual_plots_det_avg.png}} \\
\subfloat[Classification]{\includegraphics[width=0.78\linewidth]
    {figures/manual_plots_cls_whole.png}}
  \caption{Heuristic value function.}
  \label{fig:results_manual}
\end{figure}

\todo{How do the different reward functions affect performance?}

\begin{figure}[h!]
\centering
\subfloat[Detection]{\includegraphics[width=0.78\linewidth]
    {figures/entropy_plots_det_avg.png}} \\
\subfloat[Classification]{\includegraphics[width=0.78\linewidth]
    {figures/entropy_plots_cls_whole.png}}
  \caption{Entropy reward function.}
  \label{fig:results_entropy}
\end{figure}

\todo{Mention how fixedorder and fastinf learn totally different weights.}