### 2012-03-05 writeups/eccv12/eccv12.tex
final

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{ruler}
\usepackage{color}
\usepackage{subfig}
\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}
\usepackage[pagebackref=true,breaklinks=true,colorlinks,bookmarks=false]{hyperref}
\begin{document}
% \renewcommand\thelinenumber{\color[rgb]{0.2,0.5,0.8}\normalfont\sffamily\scriptsize\arabic{linenumber}\color[rgb]{0,0,0}}
% \renewcommand\makeLineNumber {\hss\thelinenumber\ \hspace{6mm} \rlap{\hskip\textwidth\ \hspace{6.5mm}\thelinenumber}}
% \linenumbers
\pagestyle{headings}

% Definitions
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\def\subsectionautorefname{section}
\definecolor{light-gray}{gray}{0.5}
\newcommand{\aside}[1]{\textcolor{light-gray}{\emph{#1}}}
\newcommand{\todo}[1]{\textcolor{red}{\emph{#1}}}
\newcommand{\cut}[1]{\textcolor{light-gray}{#1}}
\newcommand{\comment}[1]{}

\mainmatter
\def\ECCV12SubNumber{1617}  % Insert your submission number here

\title{Timely Object Recognition}

\titlerunning{ECCV-12 submission ID \ECCV12SubNumber}

\authorrunning{ECCV-12 submission ID \ECCV12SubNumber}

\author{Anonymous ECCV submission}
\institute{Paper ID \ECCV12SubNumber}

\maketitle

\begin{abstract}
In a large visual multi-class detection system, the problem of timeliness of results is crucial.
We are motivated by situations where running all detectors would take an unacceptably long time, and that the best answer must be given by some deadline.
Our system for multi-class detection aims to give the best possible results at any single point after a start time; it is terminated at a deadline time.
Toward this goal, we formulate a dynamic policy, guided by inference of the contents of the image, to decide which detector to deploy next.
We evaluate parametrizations of the policy with respect to performance in the novel AP vs. Time evaluation on the PASCAL VOC dataset.
\end{abstract}

\input{introduction.tex} % includes Evaluation section
\input{technical.tex}
\input{evaluation.tex}
\input{related.tex}
\input{future_work.tex}

\bibliographystyle{splncs}
\bibliography{sergeyk-bibtex,misc}

\end{document}