### 2012-03-05 writeups/eccv12/evaluation.tex

\begin{figure}[h!]
\centering
\subfloat[Detection]{\includegraphics[width=0.78\linewidth]
    {figures/manual_plots_det_avg.png}} \\
\subfloat[Classification]{\includegraphics[width=0.78\linewidth]
    {figures/manual_plots_cls_whole.png}}
  \caption{Heuristic value function.}
  \label{fig:results_manual}
\end{figure}

\section{Evaluation} \label{sec:evaluation}

To summarize, we evaluate our system in the multi-class, multi-label detection and classification regime.
We evaluate on the PASCAL VOC 2007 dataset: training on the training and validation data and testing on the test set.
We run our policies on all images in the test set, cutting off execution at $T_d$.

To evaluate performance at a given time, we gather all detections and the classification answers found to this point in all the recognition episodes.
As described before, we evaluate detection performance by averaging per-image performance in the multi-class regime.
We evaluate classification performance by pooling the ground truth across the dataset, in standard procedure.

We establish the baseline performance of our system by selecting actions randomly at each step.
The time setting for our evaluation is $T_s=0$, $T_d=20$ (in seconds).
As shown in Figure~\ref{fig:results_manual}, the random policy results in a roughly linear gain of performance vs. time.

To establish an upper bound on performance, we also plot the Oracle curve, obtained by re-ordering the actions with the hindsight of the results they produced.
Note that the curves turn downward in this setting: this is due to the different performance levels of the detectors: as the weaker detectors are inevitably run, they introduce more false positives and false negatives than true positives.

Figure~\ref{fig:results_manual} also shows the performance of our policy with the heuristic value function and two inference models: the MRF model and a simple fixed-order model that follows the priors.

\begin{figure}[h!]
\centering
\subfloat[Detection]{\includegraphics[width=0.78\linewidth]
    {figures/entropy_plots_det_avg.png}} \\
\subfloat[Classification]{\includegraphics[width=0.78\linewidth]
    {figures/entropy_plots_cls_whole.png}}
  \caption{Entropy reward function.}
  \label{fig:results_entropy}
\end{figure}

Figure~\ref{fig:results_entropy} shows the performance of the polices with learned weights.
A notable difference between the MRF inference model and the fixed-order model was in the weights learned: while the fixed-order model learned weights onto the bias feature only, the MRF inference model learned weights onto a mix of the entropy and probability features.