### 2012-05-31 writeups/nips12/technical.tex
paper

\section{Multi-class Recognition Policy} \label{sec:tech}
Our goal is a multi-class recognition policy $\pi$ that takes an image $\mathcal{I}$ and outputs a list of multi-class detection results by running detector and other \emph{actions} sequentially.

The policy repeatedly selects an action $a_i \in \mathcal{A}$, executes it, receiving observations $o_i$, and then selects the next action.
The set of actions $\mathcal{A}$ can include both classifiers and detectors: anything that would be useful for inferring the contents of the image.

Each action $a_i$ has an expected cost $c(a_i)$ of execution.
Depending on the setting, the cost can be defined in terms of algorithmic runtime analysis, an idealized property such as number of \emph{flops}, or simply the empirical runtime on specific hardware.
We take the empirical approach: every executed action advances $t$, the \emph{time into episode}, by its empirical runtime.

As shown in Figure~\ref{fig:evaluation}, the system is given two times: the setup time $T_s$ and deadline $T_d$.
From the setup time to the deadline, we want to obtain the best possible answer if stopped at any given time.
This corresponds to the general notion of \emph{Anytime} algorithms, and is motivated by desired flexibility in the system.

A single-number metric that corresponds to this objective is simply the ratio of the area captured under the curve to the total area between the start and deadline bounds.
We evaluate policies by this more robust metric and not simply by the final performance at deadline time for the same reason that Average Precision is used instead of a fixed Precision vs. Recall point.

\subsection{Sequential Execution}
An ``open-loop'' policy takes actions in a sequence that does not depend on observations received from previous actions.
The common classifier cascade \cite{Viola2001} is an example.
In contrast, our goal is to learn a dynamic, or ``closed-loop,'' policy, which would exploit the signal in inter-object and scene context for a maximally efficient path through the actions.

We refer to the basis for the decisions made by the decision process as the \emph{state} $s$.
The state includes the currently believed distribution over class presence variables $P(\mathbf{C}) = P(C_1, \dots, C_K)$, where we write $P(C_k)$ to mean $P(C_k=1)$

Additionally, the state records the fact that an action $a_i$ has been taken by adding it to the initially empty set $\mathcal{O}$ and recording the resulting observations $o_i$.
We refer to the current set of observations as $\mathbf{o} = \{o_i | a_i \in \mathcal{O}\}$.
The state also keeps track of the time into episode $t$, and the setup and deadline times $T_s,T_d$.

A recognition \emph{episode} takes an image $\mathcal{I}$ and proceeds from the initial state $s^0$ and action $a^0$ to the next pair $(s^1,a^1)$, and so on until $t$ exceeds $T_d$.
At that point, the policy is terminated, and a new episode can begin on a new image.

\begin{table}[h!]
\centering
\caption{Summary of the notation.}
\label{tab:notation}
\begin{tabular}{|l|l|}
  \hline
  $\mathcal{I}$ & image \\
  $C_k$         & presence of class $k \in \{1,\dots,K\}$ \\ 
  $t$           & time into episode \\ 
  $T_s$, $T_d$  & start and deadline times \\ 
  $s^j$         & belief state at step $j$ \\ 
  $\pi$         & policy function, $b \mapsto a \in \mathcal{A}$ \\
  $\mathcal{A}$ & set of actions $a$\\ 
  \comment{$\mathcal{F}$ & set of featurization actions \\}
  \comment{$\mathcal{L}$ & set of classification actions\\}
  $o_i$         & a real-valued observation upon executing $a_i \in \mathcal{A}$\\
  $\mathcal{O}$ & set of executed actions\\
  $\mathbf{o}$  & set of observations $\{o_i | a_i \in \mathcal{O}\}$\\
  $c(a_i)$        & cost of executing $a_i$, in units of $t$\\
  \hline
\end{tabular}\end{table}

\subsection{Selecting actions based on expected reward} \label{sec:value}
As our goal is to pick actions dynamically, we want a function $Q(s,a): S \times \mathcal{A} \mapsto \mathbb{R}$, where $S$ is the space of all possible states, to assign a value to a potential action, given the current state of the decision process.
We can then define the desired policy $\pi$ as simply taking the (untaken) action with the maximum value:
\begin{align}
\pi(s) = \argmax_{a_i \in \mathcal{A} \setminus \mathcal{O}} Q(s,a_i)
\end{align}

Although the action space $\mathcal{A}$ is quite manageable, consisting of the detectors and classifier we would like to run on the image, the space of possible states $B$ is infinite.
Therefore we cannot learn a tabular representation of $Q(s,a)$, and so use function approximation to represent it \cite{Sutton1998}.
We featurize the state-action pair and assume linear structure:
\begin{align}
Q^\pi(s,a_i) = \theta_\pi^\top  \phi(s,a_i)
\end{align}

The policy's performance at time $t$ is determined by the detections that are part of the set of observations $\mathbf{o}$ at the last state $s^j$ before $t$.
Detection results of unobserved classes are an empty set.
Therefore, the final AP vs. Time evaluation of the episode is a function of the history of execution $h^0=s^0,s^1,\dots,s^J$, where $J$ is the last step of the process with $t \le T_d$.
We call the final evaluation function $eval(h^0,T_s,T_d)$, and it is precisely the area under the AP vs. Time curve (normalized by total possible area) between $T_s$ and $T_d$, as determined by the detections in $\mathbf{o^j}$ for all steps $j$ in the episode $h$.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.56\linewidth]{../figures/apvst_expl.pdf}
  \caption{A per-action greedy value function that corresponds to the maximization of our objective function is the area of the horizontal slice under the curve due to the action. The figure shows this analysis for the action highlighted in orange.}
  \label{fig:rewards}
\end{figure}

As shown in \autoref{fig:rewards}, the evaluation function is additive per action, as each action can generate detections that either raise or lower the AP of all detections so far: ($\Delta ap$), and takes a certain time: $t$.
From these, we can find the area under the curve that was contributed by the action.

Specifically, as shown in Figure~\ref{fig:rewards}, we define the \emph{reward} of an action as
\begin{align}\label{eq:advanced}
R(s^j,a_i) = \frac{\Delta \text{ap}_i (t_T^j-\frac{1}{2}\Delta t_i)}{(1-\text{ap}^j)t_T^j}
\end{align}
where $t_T^j$ and $\text{ap}^j$ are the time left until deadline and the AP at state $s^j$, and $\Delta t_i$ and $\Delta \text{ap}_i$ are the time taken and AP change produced by the action $a_i$.

\todo{discuss the raw vs. the normalized area, saying we'll try both}

We can then represent the final evaluation in terms of individual rewards:
\begin{align}
eval(h^0) = \sum_{j=0}^J R(s_j,a^{j})
\end{align}

And we can represent the expected value of the final evaluation recursively in terms of the value function:
\begin{align} \label{eq:recursive_value}
Q(s^j,a_i) = \mathbb{E}_{s^{j+1}} [R(s^j,a_i) + \gamma Q(s^{j+1},\pi(s^{j+1}))]
\end{align}
where $\gamma \in [0,1]$ is a \emph{discount} value that can mitigate the effects of increasing state-transition uncertainty over long episodes.

\subsection{Learning the policy}
While we can't directly compute the expectation in \eqref{eq:recursive_value}, we can sample it by running actual episodes to gather $<s,a,r,s'>$ samples, where $r$ is the reward obtained by taking action $a$ in state $s$, and $s'$ is the following state.

\comment{
Specifically, we collect such samples to the end of the episode, and are thus able to compute the actual discounted sum of rewards for $Q(s^j,a_i)$:
\begin{align}
Q(s^j,a_i) = \theta_\pi^\top \phi(s,a_i) = \sum_{i=0}^{J-j} \lambda^i R(b_{j+i},a^{j+i})
\end{align}
}

Learning the policy is then a problem of repeatedly gathering samples with the current policy, minimizing the error between the discounted reward to the end of the episode as predicted by our current $Q(s^j,a_i)$ and the actual values gathered, and updating the policy with the resulting weights.
This is standard generalized policy iteration, or more specifically fitted Q-iteration \cite{Sutton1998,Ernst2005}.

We use $L_2$-regularized regression to minimize the error.
We run $15$ iterations of accumulating samples by running $350$ episodes, starting with a baseline policy which will be described in \autoref{sec:evaluation}, and cross-validating the regularization parameter at each iteration; samples are not thrown away between iterations.

One meta-parameter of the approach is the discount $\gamma$.
With $\gamma=0$, the value function is determined entirely by the immediate reward; learning in this case can only result in completely greedy policies.
With $\gamma=1$, the value function is determined by the expected rewards to the end of the episode, which is an approximation to the final evaluation metric.
However, the highest value for $\gamma$ is not necessarily best, due to the interplay between the expressive power of the value function and the actual state transition behavior of the world.
We experiment with several values of $\gamma$, and find a mid-level value of $0.4$ to work best.

\subsection{Feature representation}
Our policy is at its base determined by a linear function of the features of the state: $\pi(s) = \argmax_{a_i \in \mathcal{A} \setminus \mathcal{O}} \theta_\pi^\top \phi(s,a_i)$.
Since we want to be able to learn a dynamic policy, the observations $\mathbf{o}$ that are part of the state $s$ should play a role in determining the value of a potential action.
We include the following quantities as features $\phi(s,a)$:
\begin{description}
\item[$P(C_a)$] The prior probability of the class that corresponds to the detector of action $a$
\item[$P(C_0|\mathbf{o}) \ldots P(C_1|\mathbf{o})$] The probabilities of all classes, conditioned on the current set of observations.
\item[$H(C_0|\mathbf{o}) \ldots H(C_1|\mathbf{o})$] The entropies of all classes, conditioned on the current set of observations.
\end{description}

Additionally, we include the mean and maximum entropies of all classes, and time features that represent the times until start and deadline.
To formulate learning the policy as a single regression problem, we represent the features in block form, where $\phi(s,a_i)$ is a vector of size $F|\mathcal{A}|$, with $F=47$ for $20$ classes and the features as described, with all values set to $0$ except for the block corresponding to $a_i$.

Jumping ahead, we visualize the learned weights on these features in \autoref{fig:weights}, reshaped to a matrix of size $|\mathcal{A}| \times F$.

\begin{figure}[h!]
\centering
\includegraphics[width=0.87\linewidth]
    {../figures/weights.pdf}
\caption{Weights \todo{describe}}
\label{fig:weights}
\end{figure}

\subsection{Updating with observations}
\begin{figure}[h!]
\centering
\includegraphics[width=0.56\linewidth]{../figures/mrf_w_gist.pdf}
\caption{
The MRF inference model used in our system. The shaded nodes represent two actions that have already been taken; their observations have been recorded.
}
\label{fig:model}
\end{figure}

\todo{Rework from the persepctive of two potential ways of updating the beliefs with observations: just putting them in directly vs. going through MRF. De-emphasize the need for the model, treat it as an extra oomph.}

The quantities that may be useful to us for selecting which actions to deploy are the probabilities and entropies of the class presence variables $C_k$.
These allow us to look for the most probable classes given the observations.
When the policy starts, the model should present the prior distributions $P(C_k)$; as observations are accrued, the model should present the updated conditionals $P(C_k|\mathbf{o})$.

We employ a pairwise fully-connected Markov Random Field (MRF), as shown in Figure~\ref{fig:model}.
All parameters of the model are trained on fully-observed data.
Exact inference is generally intractable in this model.
Instead, we use Loopy Belief Propagation, which does not provide general convergence guarantees but has been shown to work well empirically on similar tasks \cite{Desai2009}.

The $L_i$ variables are discretized from real-valued responses of a classifier on the detections output by the deformable part-model detector we employ (see Section~\ref{sec:tech}).
The classifier is a linear SVM on the top two max detection scores in the list of detections.
The responses are discretized per variable based on the distribution of the scores on the training dataset.

The graphical model structure is set as fully-connected, but as shown in \autoref{fig:dataset_stats}, some classes are overwhelmingly unlikely to co-occurr.
Accordingly, the graph edge weights are learned with L1 regularization, which obtains the desired sparse structure \cite{Lee2006}.
 is implemented with an open-source graphical model package \cite{Jaimovich2010}.